(tentative) Rotational Spectral Unmixing of Exoplanets

### 2017-02-17 paper/main.tex
text and figure updated

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[t]
    \begin{minipage}{0.5\hsize}
    \begin{center}
\includegraphics[width=\hsize]{IGBP_simplemap.pdf}
    \end{center}
     \end{minipage}
  \begin{minipage}{0.5\hsize}
    \begin{center}
    \includegraphics[width=\hsize]{mockdata_quadrature_bandsp.pdf}
    \end{center}
\end{minipage}
  \begin{minipage}{0.5\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_quadrature_factor.pdf}
    \end{center}
 \end{minipage}
   \begin{minipage}{0.5\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_quadrature_lc.pdf}
    \end{center}
 \end{minipage}
    \caption{Our mock data based on IGBP classification map of the Earth. }
\label{fig:mockdata}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Principal Component (Hyper-)Plane\\and Graphical Conception}
\label{ss:PCplane}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Equation (\ref{eq:tilde_d_f_ast_s}) coupled with the conditions (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_sum}) indicates the geometrical relationship among ${\bf d}$, ${\bf \fast }$, and ${\bf s}$ in the $J$-th dimensional space, where $\{{\bf d}_i\}$ correspond to the points that are located on the hyperplane defined by $K$ ($<J$) number of points, $\{{\bf s}_k\} $ and are enclosed by $\{{\bf s}_k \}$. \memoYF{How is it called in one word in mathematics?}\ 
We show this graphically using the mock datasets shown in Section \ref{s:mockdata}.

The original data, however, have 4 photometric bands (i.e., $J=4$), and $J$-th dimensional space is inconvenient for visualization. 
We thus reduce the dimensionality using principal component analysis (PCA). 
Unsurprisingly we extract two dominant principal components (PCs) through PCA (other components have vertually zero contribution), which are presented in Figure \ref{fig:PCs}.

We subtract the 4-band light curves by the time averages of the respective bands, and its trajectory in the PC plane is shown in Figure \ref{fig:trajectory}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tbh!]
    \begin{center}
\includegraphics[width=\hsize]{PCA_V_jn.pdf}
    \end{center}
    \caption{Principal components.}
\label{fig:PCs}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
    \begin{center}
\includegraphics[width=\hsize]{mockdata_135deg_3types_t12_lc_noreg.png}
    \end{center}
    \caption{Trajectory on the principal component plane.}
    \label{fig:trajectory}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
    \begin{center}
\includegraphics[width=\hsize]{mockdata_135deg_3types_t12_lc_noreg.png}
    \end{center}
    \caption{Trajectory on the principal component plane.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Conversely, {\it any} set of $\{ {\bf s}_k \}$ that enclose the data points $\{{\bf d}_i\}$ in the hyperplane can be a solution of Equation (\ref{eq:tilde_d_f_ast_s}) subjected to the conditions (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_sum}), and the associated matrix, $\fast _{ik}$, can always be found. 
Additional constraint comes from the condition (\ref{eq:tilde_s_range}). This limits the range where $\{ {\bf s}_k \}$ can exist, but in general this is not sufficient to result in a unique solution. 
Therefore, predicting ${\bf \fast }$ and ${\bf s}$ from ${\bf d}$ is clearly a ill-posed problem.

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
    \begin{center}
	\includegraphics[width=\hsize]{mockdata_3types_map.pdf}
    \includegraphics[width=\hsize]{mockdata_3types_albd.pdf}
	\includegraphics[width=\hsize]{mockdata_90deg_3types_t360_lc.pdf}
    \end{center}
    \caption{Top: our mock data based on IGBP classification map of the Earth, with 3 surface types (ocean: blue, soil: red, vegetation: green). }
\label{fig:mockdata}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Graphical Conception on PC (Hyper-)Plane\\and Formal Degeneracy}
\label{ss:PCplane}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Equation (\ref{eq:tilde_d_f_ast_s}) coupled with the conditions (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_sum}) indicates the geometrical relationship among ${\bf d}$, ${\bf \fast }$, and ${\bf s}$ in the $J$-th dimensional space, where $\{{\bf d}_i\}$ correspond to the points that are located on the hyperplane defined by $K$ ($<J$) number of points, $\{{\bf s}_k\} $ and are enclosed by $\{{\bf s}_k \}$. \memoYF{How is it called in one word in mathematics?}\ 
We show this graphically below, using the mock datasets shown in Section \ref{s:mockdata}.

The original data, however, have 4 photometric bands (i.e., $J=4$), and $J$-the dimensional space is inconvenient for visualization. 
We thus reduce the dimensionality using principal component analysis (PCA). 
As mentioned in \citet{Cowan2011}, the number of the spectrally distinct surface types is number of PCs plus 1 in theory. 
Reasonably, we extract two dominant principal components (PCs) through PCA (other components have virtually zero contribution) from the mock light curves. 
The PCs extracted from the light curves at $\alpha = 90^{\circ }$ (olive lines in Figure \ref{fig:mockdata}) are presented in Figure \ref{fig:PCs}. 
Note that, although the PCs from other light curves are not necessarily same as the one shown in Figure \ref{fig:PCs}, the plane they make should be the same plane (under the noiseless condition). Thus, in the following, we adopt these two PCs as the axes of the PC plane.

Figure \ref{fig:trajectory} shows the trajectories of the mock light curves with varying phase angle. 
The excursion of the trajectory is larger for the light curve at a larger phase angle, i.e., at crescent phase, and shrinks as the phase angle decreases when larger surface area are averaged. 
The points in the figure indicate the input albedo spectra of the three surface types. 
Because the apparent albedo at any time of the observations is a linear combination of input surface colors (Equation \ref{eq:tilde_d_f_ast_s}) under the conditions (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_sum}), the trajectories should always be inside the triangle defined by these points.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tbh!]
    \begin{center}
\includegraphics[width=\hsize]{PCA_V_jn.pdf}
    \end{center}
    \caption{Principal components of the light curves shown in the bottom of Figure \ref{fig:mockdata}. }
\label{fig:PCs}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
    \begin{center}
\includegraphics[width=\hsize]{mockdata_3types_t360_lc_PCplane.pdf}
    \end{center}
    \caption{Lines: trajectory of the 4-band light curves projected on the PC plane, in the case of $\alpha = 135^{\circ }$ (thick indigo), $\alpha = 90^{\circ }$ (olive), and $\alpha = 45^{\circ }$ (gold). Points: The input albedo spectra of ocean (blue), soil (red), and vegetation (green), projected on the PC plane. The surface albedo spectra cannot be located in the ray region, because it violates the condition (\ref{eq:tilde_s_range}). }
    \label{fig:trajectory}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Degeneracy}
%\label{ss:degeneracy}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

When it comes to the point where we must estimate the points given the trajectory/-ies, however, {\it any} set of $\{ {\bf s}_k \}$ that enclose the data points $\{{\bf d}_i\}$ in the hyperplane can be a solution of Equation (\ref{eq:tilde_d_f_ast_s}) subjected to the conditions (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_sum}). 
The associated matrix, $\fast _{ik}$, can always be found. 
Additional constraint comes from the condition (\ref{eq:tilde_s_range}). This limits the range where $\{ {\bf s}_k \}$ can exist, but in general this is not sufficient to result in a unique solution. 
Therefore, predicting ${\bf \fast }$ and ${\bf s}$ from ${\bf d}$ is clearly a ill-posed problem.

@@

Given that any combinations that enclose the trajectory/-ies can be a solution, the choice of the solution critically depends on the prior probability distribution. 
% Thus, we need to be careful about the implicit assumptions that are made  
With no assumptions or information on the spectral albedo of the surface types, it may be reasonable to assume that any points in the PC plane except for the inhibited regions are equally likely to correspond to a surface type. 
While such an assumption does not seem to work at all to choose the surface types, it does put some constrains. 
This is because of the relative configuration of the trajectory and the permitted region. 
For example, in Figure \ref{fig:trajectory}, in order to enclose the trajectory with three surface types within the permitted region, we must have at least one point near the left bottom corner of the permitted region (white triangle); this is consistent with the fact that one of the input albedo spectra (ocean) resides there.

In order to see this more quantitatively, we made a grid on the PC plane  with an interval of ?? (shown in Figure \ref{fig:gridsearch}) and considered all combinations of three grid points. 
Then, we assumed that all of the combinations that enclose the trajectory (in the case of $\alpha = 1356{\circ }$) are equally likely to be a solution, and find the marginalized probability of having a surface type at each location of the PC plane, as shown in Figure \ref{fig:gridsearch}. 
%
As expected, we find a relatively definite peak near the point of ocean. 
On the other hand, the peak that corresponds to soil is blurred.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Constraining the Surface Types}
%\label{ss:constraining}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
